---
title: "2019 Instructor Demographics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```


```{r include=FALSE}
by_gender_country <- readr::read_csv("https://raw.githubusercontent.com/carpentries/equity-and-inclusion/master/data/instructor-demographics-gender-country_2019_06_17.csv")

by_domain <- readr::read_csv("https://raw.githubusercontent.com/carpentries/equity-and-inclusion/master/data/instructor-demographics-domain_2019_06_17.csv")
```

```{r include=FALSE}
# Function that makes a table of counts and percentages
# question_n is set by default to the number of respondents in the survey. This value may have to be set by question.
question_n <- nrow(by_gender_country)

tally_and_perc <- function(df, colname, na.rm = FALSE, question_n){
  quo_colname <- enquo(colname)

  df %>%
    group_by(!!quo_colname) %>%
    tally() %>%
    filter(if_else(rep(na.rm, nrow(.)),
                  !is.na(!!quo_colname),
                  as.logical(rep(1, nrow(.))))) %>%
    mutate(`%` = round(n / question_n * 100, 1))
}

# function to compute number of non-NA responses to a question

n_responses_to_the_question <- function(df, from_colname, to_colname) {

  quo_from_colname <- enquo(from_colname)
  quo_to_colname <- enquo(to_colname)

  rowsums <-
df %>%
  select(UQ(quo_from_colname):UQ(quo_to_colname)) %>%
  # check that each row has a value for at least one col
  # convert to numeric, if NA, then put a zero, otherwise 1
  map_df(., ~ifelse(is.na(.x), 0, 1)) %>%
  # sum to see if any rows have no reponses
  rowSums()

# for all behaviors
idx <- ifelse(rowsums == 0, FALSE, TRUE)
sum(idx)
}
```

## Certified Instructors by Country
```{r}
## Tally Certified Instructors by Country

instructors_by_country <- by_gender_country %>%
    group_by(country) %>%
    tally(sort = TRUE) %>%
    mutate(percent = round(100 * (n/sum(n)), 1)) %>% # add the % col
    filter(!is.na(country))

get_instructors_by_country_percent <- function(country_var, .data = instructors_by_country) {
    extract_percentage(.data, .var = country, .value = country_var)
}

instructors_by_country
```

## Certified Instructors by Gender*

We recognize that our practices for collecting inclusive gender data for our instructors is an area of improvement. Please be patient with us as we are making plans to improve our data collection practices.

```{r}
## Tally Certified Instructors by Gender

instructors_by_gender <- by_gender_country %>%
    group_by(gender) %>%
    tally(sort = TRUE) %>%
    mutate(percent = round(100 * (n/sum(n)), 1)) %>% # add the % col
    filter(!is.na(gender))

get_instructors_by_gender_percent <- function(gender_var, .data = instructors_by_gender) {
    extract_percentage(.data, .var = gender, .value = gender_var)
}


instructors_by_gender
```



## Certified Instructors by Domain

Domain may include multiple rows for each badged Instructor. Instructors who have more than one domain of expertise are listed one time for each domain. The person_id variable can be used to get a count of total number of Instructors represented, though this will sum to more than 100%.

```{r}
## Tally Certified Instructors by Domain

instructors_by_domain <- by_domain %>%
    group_by(domain) %>%
    tally(sort = TRUE) %>%
    mutate(percent = round(100 * (n/sum(n)), 1)) %>% # add the % col
    filter(!is.na(domain))

get_instructors_by_domain_percent <- function(domain_var, .data = instructors_by_domain) {
    extract_percentage(.data, .var = domain, .value = domain_var)
}

instructors_by_domain
```

